ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-07-20 pyked.tex
Update examples to v0.1.6 format

This section will provide a brief introduction to the current version of the
\ck{} database format, v0.1.6, which represents a functional standard
for documenting autoignition measurements in shock tubes and rapid compression machines.
This is not a complete listing of available fields and more complete examples of
\ck{} files can be found in the \pybox|tests| directory of the \pk{} package. In
addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
files. We indicate \yabox|yaml| keywords or values using text with gray background,
and \pybox|Python| code using text with a yellow background.

@@

The \yabox|value| can be a string, integer, or floating point number. The \ck{}
format is designed to include all of the information necessary to simulate a
given experiment and can be broken into two main sections. The first section
details all of the ``meta'' information about the experiment and the \ck{} file.
In this section, the common \yabox|author| mapping is used for any fields where
an author must be specified. The \yabox|author| mapping contains the following
fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yabox|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|journal| (required, string): The name of the publishing
        journal
        \item \yabox|year| (required, integer): The year of publication
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|volume| (required, integer): The journal volume number
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
    \end{itemize}
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|apparatus| (required, string): The type of apparatus used to
    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
\end{itemize}

@@

The \ck{} file that will be used in this example can be found in the
\pybox|tests| directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: (1, 0)
chemked-version: 0.0.1
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, there is just a single datapoint in this file, which describes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure, compression time, ignition type, and the volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
  - temperature: 297.4 kelvin
    ignition-delay: 1.0 ms
    pressure: 958.0 torr
    composition:
      - species-name: H2
        InChI: 1S/H2/h1H
        mole-fraction: 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        mole-fraction: 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        mole-fraction: 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        mole-fraction: 0.63125
    ignition-type:
      target: pressure
      type: d/dt max
    compression-time: 38.0 ms
    volume-history:
      time:
        units: s
        column: 0
      volume:
        units: cm3
        column: 1
      values:
        - [0.00E+000, 5.47669375000E+002]
        - [1.00E-003, 5.46608789894E+002]
        - [2.00E-003, 5.43427034574E+002]
        - [3.00E-003, 5.38124109043E+002]
        ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. A user
might want to load the information from this file to perform a simulation using
Cantera to calculate the ignition delay. All of the information required to
perform this simulation is present in the \ck{} file, with the exception of a
chemical kinetic model for \ce{H2}\slash {CO} combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

@@

The \ck{} file used in this example can be found in the \pybox|tests| directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    - species-name: t-butanol
      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
      mole-fraction: 0.003333333
    - species-name: O2
      InChI:  1S/O2/c1-2
      mole-fraction: 0.04
    - species-name: Ar
      InChI:  1S/Ar
      mole-fraction: 0.956666667
  ignition-type:  &ign
    target: OH* concentration
    type: 1/2 max
\end{yamlbox}
%
This block uses the ability of YAML files to define an anchor with the
\yabox|&| symbol and refer to that section later with the \yabox|*| symbol, as
can be seen in the definition of the \yabox|datapoints| section:
%
\begin{yamlbox}
datapoints:
  - temperature: 1459 kelvin
    ignition-delay: 347 us
    pressure: 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1389 kelvin
    ignition-delay: 756 us
    pressure: 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1497 kelvin
    ignition-delay: 212 us
    pressure: 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1562 kelvin
    ignition-delay: 105 us
    pressure: 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}
The \yabox|common-properties| section is not required, but can save space and
help avoid errors when all data points share some common values.

>>>>

This section will provide a brief introduction to the current version of the
\ck{} database format, v0.1.6, which represents a functional standard
for documenting autoignition measurements in shock tubes and rapid compression machines.
This is not a complete listing of available fields and a complete reference for the
schema can be found in the online documentation.
\footnote{\url{https://pr-omethe-us.github.io/PyKED/schema-docs.html}}
In addition, in Section~\ref{sec:usage-example}, we will provide examples of \ck{}
files. We indicate \yabox|yaml| keywords or values using text with gray background,
and \pybox|Python| code using text with a yellow background.

@@

The \yabox|value| can be a string, integer, or floating point number. The \ck{}
format is designed to include all of the information necessary to simulate a
given experiment. A \ck{} file can generally be broken into two main sections.
The first section details all of the ``meta'' information about the experiment
and the \ck{} file. In this section, the common \yabox|author| mapping is used
for any fields where an author must be specified. The \yabox|author| mapping
contains the following fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's ORCID number
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema that this file conforms to
    \item \yabox|reference| (required, mapping): The reference information for
    the article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|journal| (required, string): The name of the publishing
        journal
        \item \yabox|year| (required, integer): The year of publication
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|volume| (required, integer): The journal volume number
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
    \end{itemize}
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|apparatus| (required, string): The type of apparatus used to
    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
\end{itemize}

@@

The \ck{} file that will be used in this example can be found in the
\pybox|tests| directory of the PyKED
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: 0
chemked-version: 0.1.6
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the dataset comes from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, there is just a single datapoint in this file, which describes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure, compression time, ignition type, and the volume history that specifies
how the volume of the reactor varies with time, for simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
- temperature:
    - 297.4 kelvin
  ignition-delay:
    - 1.0 ms
  pressure:
    - 958.0 torr
  composition:
    kind: mole fraction
    species:
      - species-name: H2
        InChI: 1S/H2/h1H
        amount:
          - 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        amount:
          - 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        amount:
          - 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        amount:
          - 0.63125
  ignition-type:
    target: pressure
    type: d/dt max
  compression-time:
    - 38.0 ms
  volume-history:
    time:
      units: s
      column: 0
    volume:
      units: cm3
      column: 1
    values:
      - [0.00E+000, 5.47669375000E+002]
      - [1.00E-003, 5.46608789894E+002]
      - [2.00E-003, 5.43427034574E+002]
      ...
\end{yamlbox}
%
The values for the \yabox|volume-history| are truncated here to save space. A user
might want to load the information from this file to perform a simulation using
Cantera to calculate the ignition delay. All of the information required to
perform this simulation is present in the \ck{} file, with the exception of a
chemical kinetic model for \ce{H2}\slash {CO} combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

@@

The \ck{} file used in this example can be found in the \pybox|tests| directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    kind: mole fraction
    species:
      - species-name: t-butanol
        InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
        amount:
          - 0.003333333
      - species-name: O2
        InChI:  1S/O2/c1-2
        amount:
          - 0.04
      - species-name: Ar
        InChI:  1S/Ar
        amount:
          - 0.956666667
  ignition-type: &ign
    target: OH*
    type: 1/2 max
\end{yamlbox}
%
This block uses the ability of YAML files to define an anchor with the
\yabox|&| symbol and refer to that section later with the \yabox|*| symbol, as
can be seen in the definition of the \yabox|datapoints| section:
%
\begin{yamlbox}
datapoints:
  - temperature:
      - 1459 kelvin
    ignition-delay:
      - 347 us
    pressure:
      - 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1389 kelvin
    ignition-delay:
      - 756 us
    pressure:
      - 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1497 kelvin
    ignition-delay:
      - 212 us
    pressure:
      - 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1562 kelvin
    ignition-delay:
      - 105 us
    pressure:
      - 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}
The \yabox|common-properties| section is not required, but can save space and
help avoid errors when several data points share some common values. However, even if
properties are specified in the \yabox|common-properties| section, they must also be
explicitly specified in the \yabox|datapoints| section for each data point, as can be
seen in the previous code.
